Unsupervised embeddings of organizations capture latent structure of scientific mobility

### 2020-03-03 paper/current/main.tex
revise introduction para

\title{Unsupervised embeddings of institutions capture latent structure of scientific mobility} %{{{
\date{\today}
\maketitle %}}}

@@

Scientific mobility drives innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, jonkers2013return, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige}. 
Common models of mobility, such as the Gravity Model~\autocite{curiel2018citygravity, lewer2008immigrationgravity, jung2008highwaygravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity, simini2012universal}, describes the of mobility using \textit{distance} between places.
However, amid a globalizing scientific enterprise~\autocite{czaika2018globalisation}, geographic distance is losing its relevance for understanding mobility, whereas other abstract notions of distance~\autocite{boschma2005proximity} pose challenges; too many choices.
Accordingly, the complexity of factors driving global scientific mobility~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics} necessitates simplified models for reasoning about and crafting effective policy. 
Here, we demonstrate a data-driven approach of learning an effective distance that encodes the latent structure of mobility using a neural embedding model~\autocite{mikolov2013word2vec} by sourcing publications of disambiguated authors from the Web of Science database between 2008 and 2019. 
Distance between vectors in the learned embedding space explain and predict mobility between organizations better than  geographic distance or network distance. We argue that such power emerges from an underlying connection between embeddings and the gravity model.
In addition to geography, these embeddings also capture latent religious, cultural, social, and linguistic, and even prestige of organizations at multiple levels of granularity.
With this approach, we can learn scalable, dense, and meaningful vector representations of scientific mobility directly from the data and it can be generalized methodology of reasoning about all kinds of mobility.
\end{abstract}

@@

When scientists move, so does their knowledge, and for scientific mobility, like all human movement, distance matters. 
However, how exactly do we measure the distance between two places?
Australia may be physically close to Indonesia, but in terms of language, history, and shared culture, they are essentially neighbors to geographically distant countries like the United Kingdom or United States. 
There are many ways to think about the distance~\autocite{boschma2005proximity}, but only geographic distance can be meaningfully quantified. 
In turn, models of human mobility, such as the Gravity~\autocite{zipf1946gravity} and Radiation models ~\autocite{simini2012universal}, rely almost exclusively on geographic distance, rather than other less measurable though important forms. 
Encoding distance in a way that considers not just geography, but also historical, linguistic, cultural, and cognitive proximity would improve our models and thus ability to reason about science and immigration policies.

The skip-gram negative-sampling embedding (\SGNS) model preserves semantic relationships between words in a dense, continuous and meaningful low-dimensional vector representation~\autocite{mikolov2013word2vec}, and just as its has been applied to and revealed insights implicit in text~\autocite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry, hamilton2016diachronic}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}, so too might this model encode the latent structures of global scientific mobility. 
As scientist's careers progress, they move institutions or pick up additional affiliations, forming co-affiliation or \textit{mobility trajectories} that can be derived from their publication histories~\autocite{robinson2019mobility, sugimoto2017mostimpact}. 
We embed mobility trajectories by treating organizations analogously to words and trajectories analogously to sentences, and feed the resulting "sentence" into the standard \SGNS{} model (Fig.~\ref{fig:methods:mobility_sentence}).
The result is a vector representation of organizations in which the \textit{cosine distance} between vectors represents the tendency for pairs of organizations to occur near one another in the same career path. 
In addition to encoding latent structure, this approach is computationally efficient, scalable for large dataset, robust to noise, and allow for a range of unique analyses that leverage their semantic properties using simple vector arithmetic.

@@

We embed the trajectories of the three million name-disambiguated authors who were \textit{mobile}---having more than one affiliation---between 2008 and 2019 as evidenced by their publications indexed in the Web of Science databse. 
We adopt the Gravity Model of mobility~\autocite{zipf1946gravity}, one of the most common and broadly applied mobility models~\autocite{curiel2018citygravity, lewer2008immigrationgravity, jung2008highwaygravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity} to assess how well the embedding captures mobility. After fitting the parameters of the gravity model, we derive the \textit{expected flux} $\widetilde{f_{ij}}$ from the obtained gravity model, defined as,

Where the flux, or number of authors co-affiliated with two organizations during the time period $f_{ij}$, increases with the organization's sizes, $m_{i}$ and $m_{j}$, and decays as a function of their distance, $f(r_{ij})$. 
We consider two distances, the geographic distance between the organizations as a baseline, and the \textit{embedding distance} defined as the cosine distance between the organization's vector in the trained embedding space. 
We argue that the performance of a distance measure can be observed from their ability to explain expected mobility between organizations, as well as to predict flux using the gravity model. 
For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function. selected as the best performing for each case.

>>>>

\title{Unsupervised embeddings of organizations capture latent structure of scientific mobility} %{{{
\date{\today}
\maketitle %}}}

@@

Scientific mobility drives innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, jonkers2013return, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity, simini2012universal}, describes the of mobility using \textit{distance} between places.
However, amid a globalizing scientific enterprise~\autocite{czaika2018globalisation}, geographic distance is losing its relevance for understanding mobility, whereas other abstract notions of distance~\autocite{boschma2005proximity} pose challenges; too many choices.
Accordingly, the complexity of factors driving global scientific mobility~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics} necessitates simplified models for reasoning about and crafting effective policy. 
Here, we demonstrate a data-driven approach of learning an effective distance that encodes the latent structure of mobility using a neural embedding model~\autocite{mikolov2013word2vec} by sourcing publications of disambiguated authors from the Web of Science database between 2008 and 2019. 
Distance between vectors in the learned embedding space explain and predict mobility between organizations better than  geographic distance or network distance. We argue that such power emerges from an underlying connection between embeddings and the gravity model.
In addition to geography, these embeddings also capture latent religious, cultural, social, and linguistic, and even prestige of organizations at multiple levels of granularity.
With this approach, we can learn scalable, dense, and meaningful vector representations of scientific mobility directly from the data and it can be generalized methodology of reasoning about all kinds of mobility.
\end{abstract}

@@

When scientists move, so does their knowledge, and for scientific mobility, like all human movement, distance matters. 
However, how exactly do we measure the distance between two places?
Although many alternative ways to quantify the distance~\autocite{boschma2005proximity} exist, geographic distance is an unambiguous and foundational quantity for the distance.
In turn, models of human mobility, such as the Gravity~\autocite{zipf1946gravity} and Radiation models ~\autocite{simini2012universal}, rely almost exclusively on geographic distance, yet, geographic distance fails to capture historical, linguistic, cultural, and cognitive proximity.
For instance, Australia may be physically close to Indonesia, but in terms of language, history, and shared culture, they are essentially neighbors to geographically distant countries like the United Kingdom.
Then, how can we quantify meaningful, useful distance between institutions, cities and countries?

% Let's rephease and reuse to somewhere else:)
%Encoding distance in a way that considers not just geography, but also historical, linguistic, cultural, and cognitive proximity would improve our models and thus ability to reason about science and immigration policies.

It has been demonstrated that neural embedding model preserves semantic relationships between words in a dense, continuous and meaningful low-dimensional vector representation~\autocite{mikolov2013word2vec}, and just as its has been applied to and revealed insights implicit in text~\autocite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry, hamilton2016diachronic}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}. In addition to encoding latent structure, this approach is computationally efficient, scalable for large dataset, robust to noise, and allow for a range of unique analyses that leverage their semantic properties using simple vector arithmetic. Here, we propose a neural embedding models an effective way to estimate distance between  from massive trajectory data.

As scientist's careers progress, they move organizations or pick up additional affiliations, forming co-affiliation or \textit{mobility trajectories} that can be derived from their publication histories~\autocite{robinson2019mobility, sugimoto2017mostimpact}. 
From these mobility trajectories, by treating geo-located organizations analogously to words and trajectories analogously to sentences, we obtain vector-space embeddings of organizations  from neural embedding method, which learn vector representation of organizations by using neural networks (Fig.~\ref{fig:methods:mobility_sentence}).
The result vector of organizations is representation that the \textit{cosine distance} between vectors represents the distance in the landscape of the scientific mobility data.

@@

We use the trajectories of the three million name-disambiguated authors who were \textit{mobile}---having more than one affiliation---between 2008 and 2019, as evidenced by their publications indexed in the Web of Science databse. 
We adopt the Gravity Model of mobility~\autocite{zipf1946gravity, barbosa2018human}, one of the most common and broadly applied mobility models~\autocite{curiel2018citygravity, lewer2008immigrationgravity, jung2008highwaygravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity} to assess how well the embedding captures mobility. After fitting the parameters of the gravity model, we derive the \textit{expected flux} $\hat{t_{ij}}$ from the obtained gravity model, defined as,

\begin{equation}
	\label{eq:gravity_basic}
	\hat{t_{ij}} = Cm_{i}m_{j}f(r_{ij})
\end{equation}

Where the flux, or number of authors co-affiliated with two organizations during the time period $t_{ij}$, increases with the organization's sizes, $m_{i}$ and $m_{j}$, and decays as a function of their distance, $f(r_{ij})$. 
We consider two distances, the geographic distance between the organizations as a baseline, and the \textit{embedding distance} defined as the cosine distance between the organization's vector in the trained embedding space. 
We argue that the performance of a distance measure can be observed from their ability to explain expected mobility between organizations, as well as to predict flux using the gravity model. 
For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function. selected as the best performing for each case.
